FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2018-01-09 ECOOP18/sections/Auxiliary.tex
Fix review comments
Proof of Determinacy in latex

\subsection{Method Lookup Algorithm in \mbody{}}\label{subsec:mbodydef}

$\mbody(m, I_d, I_s)$ denotes the method body lookup function.
We use $I_d, I_s$, since $\mbody$ is usually invoked by a receiver of a method $m$, with its dynamic type $I_d$ and static type $I_s$. Such a function returns the most specific method implementation. More
accurately, $\mbody$ returns the parameters, returned expression
(empty for abstract methods) and the types for the method. It considers both originally-defined methods and hierarchical overriding methods, so $\mostSpecific$ and $\mostSpecificOverride$ (see the definition in Section~\ref{sec:mostSpecific} and Section~\ref{sec:mostSpecificOverride}) are both invoked.
 The formal definition gives the expected results
for the earlier examples in Figure~\ref{fig:examplesmbody}.

@@

\begin{theorem}[Type Soundness]~\label{theorem_soundness}
If $\judgeewf \o {e : I}$ and $e \to^* e'$ with $e'$ a normal form, then $e'$ is 
a value $v$ with $\judgeewf \o {v:I}$.
\end{theorem}
\begin{proof}
Immediate from Theorem~\ref{theorem_subject} and Theorem~\ref{theorem_progress}.
\end{proof}
Note that in Theorem~\ref{theorem_progress}, ``$\#(\overline{x})$'' denotes the length of
$\overline{x}$.

>>>>

\subsection{The Method Lookup Algorithm in \mbody{}}\label{subsec:mbodydef}
$\mbody(m, I_d, I_s)$ denotes the method body lookup function.
We use $I_d, I_s$, since $\mbody$ is usually invoked by a receiver of a method $m$, with its dynamic type $I_d$ and static type $I_s$. Such a function returns the most specific method implementation. More
accurately, $\mbody$ returns the parameters, returned expression
(empty for abstract methods) and the types for the method. It considers both originally-defined methods and hierarchical overriding methods, so $\mostSpecific$ and $\mostSpecificOverride$ (see the definition in Section~\ref{sec:mostSpecific} and Section~\ref{sec:mostSpecificOverride}) are both invoked.
 The formal definition gives the expected results
for the earlier examples in Figure~\ref{fig:examplesmbody}.

@@

\begin{theorem}[Type Soundness]~\label{theorem_soundness}
If $\judgeewf \o {e : I}$ and $e \to^* e'$ with $e'$ a normal form, then $e'$ is 
a value $v$ with $\judgeewf \o {v:I}$.
\end{theorem}
\begin{proof}
Immediate from Theorem~\ref{theorem_subject} and Theorem~\ref{theorem_progress}.
\end{proof}

\begin{theorem}[Determinacy of One-Step Evaluation]~\label{theorem_determinacy}
	If $t \to t'$ and $t \to t''$, then $t' = t''$.
\end{theorem}
\begin{proof}
See Appendix~\ref{appendix_proof}.
\end{proof}

Note that in Theorem~\ref{theorem_progress}, ``$\#(\overline{x})$'' denotes the length of
$\overline{x}$.

### 2018-01-09 ECOOP18/sections/TypeSafety.tex
Proof of Determinacy in latex

\subsubsection{Proof for Theorem~\ref{theorem_progress}}
\begin{proof}

Since $e$ is well-typed, by \textsc{(T-Invk)} and \textsc{(T-Anno)} we know that
$$I \subtype J\textrm{, and }\mbody(m, J, J) \textrm{ is defined}$$

@@

By \textsc{(T-New)}, $\canOverride(I) = True$. By the definition of $\canOverride$, any $J_0\in\mostSpecific(m, I, I)$ satisfies that
$\mostSpecificOverride(m, I, J_0)$ contains only one interface, in which the $m$ that overrides $J_0$ is a concrete method. Therefore $\mbody(m, I, J)$ also provides a concrete method, which finishes the proof.
\end{proof}

>>>>

\subsubsection{Proof for Theorem~\ref{theorem_progress}}
\begin{proof}
Since $e$ is well-typed, by \textsc{(T-Invk)} and \textsc{(T-Anno)} we know that
$$I \subtype J\textrm{, and }\mbody(m, J, J) \textrm{ is defined}$$

@@

By \textsc{(T-New)}, $\canInstantiate(I) = True$. By the definition of $\canInstantiate$, any $J_0\in\mostSpecific(m, I, I)$ satisfies that
$\mostSpecificOverride(m, I, J_0)$ contains only one interface, in which the $m$ that overrides $J_0$ is a concrete method. Therefore $\mbody(m, I, J)$ also provides a concrete method, which finishes the proof.
\end{proof}

\subsubsection{Proof for Theorem~\ref{theorem_determinacy}}
\begin{proof}	
The Proof is done by induction on a derivation of $t \to t'$, following the book  \emph{TAPL}.
\begin{itemize}
	\item If the last rule used in the derivation of $t \to t'$ is \textsc{(S-Invk)}, then we know that $t$ has the form $\left((J)\new{I}\right).m(\overline{v})$ with $I, J, m$ determined. Now it is obvious that the last rule in the derivation of $t \to t''$ should also be \textsc{(S-Invk)} with the same $I, J, m$. Since $\mbody(m, I, J)$ is a \emph{function} that given the same input will calculate the same result, we know the two induction results are the same, thus $t' = t''$ is immediately proved.

	\item If the last rule used in the derivation of $t \to t'$ is \textsc{(C-Receiver)}, then $t$ has the form $e_0.m(\overline{e})$ and $e_0 \to e_0'$. Since $e_0$ is not a value, the last rule used in $t \to t''$ has to be \textsc{(C-Receiver)} (other rules do not match) too. Assume in the reduction $t \to t''$, $e_0 \to e_0''$, thus $e_0'.m(\overline{e}) = e_0''.m(\overline{e})$. Thus, $t' = t''$ proved.

	\item If the last rule used in the derivation of $t \to t'$ is \textsc{(C-StaticType)}, then $t$ is fixed to be $\new{I}$. The last rule used in $t \to t''$  has to be \textsc{(C-StaticType)}, and obviously, $t' = t'' = (I)\new{I}$.

	\item If the last rule used in the derivation of $t \to t'$ is \textsc{(C-FReduce)}, then $t$ has the form $(I)e$ and $e \to e'$. The last rule used in $t \to t''$ cannot be \textsc{(C-StaticType)} because it requires $t$ to be $\new{I}$; it can neither be \textsc{(C-AnnoReduce)} because it requires $t$ to be $(I)((J)\new{K})$ where $(J)\new{K}$ is already a value. So the last rule used in $t \to t''$ can only be \textsc{(C-FReduce)} (other rules do not match). Assume in the reduction $t \to t''$, $e \to e''$, and $(I)e \to (I)e''$. By induction hypothesis, $e' = e''$, thus $t' = t''$ proved.

	\item If the last rule used in the derivation of $t \to t'$ is \textsc{(C-AnnoReduce)}, then the form of $t$ is fixed to be $(I)((J)\new{K})$. Since $(I)((J)\new{K})$ is not reducible, the rule \textsc{(C-FReduce)} does not apply. The only rule applies in $t \to t''$ is \textsc{(C-AnnoReduce)}. Thus $t' = t'' = (I)\new{K}$ proved.

	\item If the last rule used in the derivation of $t \to t'$ is \textsc{(C-Args)}, then $t$ has the form $v.m(..., e, ...)$ and $e \to e'$. The last rule used in $t \to t''$ cannot be \textsc{(S-Invk)} because it requires all arguments to be values. Thus only \textsc{(C-Args)} applies to $t \to t''$. Assume in the reduction $t \to t''$, $e \to e''$. By induction hypothesis, $e' = e''$, thus $v.m(..., e', ...) = v.m(..., e'', ...)$, thus $t' = t''$ proved.
\end{itemize}
\end{proof}

\begin{comment}
If the last rule used in the derivation of t  → t′ is E-IfTrue, then we know that t has the form if t1 then t2 else t3, where t1 = true. But now it is obvious that the last rule in the derivation of t  → t′′ cannot be E-IfFalse, since we cannot have both t1 = true and t1 = false. Moreover, the last rule in the second derivation cannot be E-If either, since the premise of this rule demands that t1  → t′1 for some t′1, but we have already observed that true does not evaluate to anything. So the last rule in the second derivation can only be E-IfTrue, and it immediately follows that t′ = t′′.
\end{comment}
